Reproducible Research Practices and Barriers to Reproducible Research in Geography: Insights from a Survey

### 2023-01-23 docs/manuscript/manuscript.tex

%%%%%%%%%%%
\subsubsection*{Reproducible Research Practices and Barriers to Reproducibility}
Geographic researchers are on average familiar with multiple reproducible research practices ($\overline{f}=3.26$), but reported limited use of these practices in their own work ($\overline{p}=1.44$).
More than half of all researchers reported familiarity with with data sharing (86.7\%), open source software (85.3\%), field and lab notebooks (67.0\%), and code sharing (59.2\% ).
However, while familiar with these practice, a far smaller number of researchers reported using these practices regularly in their own work. 
Less than half of the researchers surveyed reported sharing their data (44.5\%), using open source software (38.1\%), or using field or lab notebooks to record their work (40.0\%), or sharing their code (18.8\%) most or all of the time. 
Only a small subset of researchers reported familiarity with the pre-registration of research designs and protocols (27.5\%), or regular use of this practice (2.7\%).

Researcher familiarity and use of reproducible research practices also varied by disciplinary subfield and methodological approach. 
Researchers that identified as physical geographers ($\overline{f_{PH}}=3.71$) or methodologists and GIScientists ($\overline{f_{MT}}=3.97$) reported being familiar with more reproducible researcher practices than human  geographers ($\overline{f_{HU}}=2.35$) and those focused on nature and society ($\overline{f_{NS}}=2.28$).
Researcher practices similarly diverged by subfield, but no subset of researchers reported using on average more than two of these practices regularly in their work.  
When grouped by methodological approach, Quantitative and mixed-methods researchers reported substantially greater familiarity ($\overline{f_{QN}}=3.75$, $\overline{f_{MX}}=3.56$) and use ($\overline{p_{QN}}=1.63$, $\overline{p_{MX}}=1.69$) of reproducible researcher practices than qualitative researchers ($\overline{f_{QL}}=1.75$, $\overline{p_{QL}}=0.48$).

Differences in practice familiarity and use across subfields and approaches were greatest for practices more typical of quantitative workflows. 
For example, quantitiative and mixed-methods researchers reported familiarity with open source software (QNT=96.7\%, MIX=89.5\%) and code sharing (QNT=81.5\%, MIX=55.7\%).
In contrast, a smaller percentage of qualitative researchers reported familiarity with these same practices (OS=52.5\%, Code=12.5\%).

Even among quantitiative and mixed-methods researchers familiarity with reproducible research practices did not translate into regular use of those practices.
For example, just over half of quantitative (51.2\%) and mixed-methods (51.8\%) researchers reported sharing their data most of the time.
Less than a third of these researchers reported sharing their code most of the time. 
These rates were lower for qualitative geographers, of which only a small subset reported using open source software (10.0\%) or field notes (20.0\%), or sharing data (12.5\%) and code (5.0\%) most of the time.

\begin{itemize}
    \item Use Q9 to contextualize the frequency measure.
    \item Use details from Q14 to enrich the discussion of barriers
\end{itemize}

@@

@@

Start with the outline of what the population is here because not all respondents answered these questions. Start with descriptives that show this sub-sample of respondents. Summary discussion of Q10-Q12.

>>>>

@@

%%%%%%%%%%%
\subsubsection*{Familiarity and use of Reproducible Research Practices}
The reproducibility of research is on the minds of geographic researchers. 
The majority of survey respondents reported thinking about the reproducibility of their own research (80.7\%), discussing reproducibility with a colleague (70.6\%), and questioning the reproducibility of published work (57.3\%) in the past two years. 
More than half of the researchers we surveyed (52.8\%) also reported considering reproducibility while peer reviewing a grant proposal or publication during the same time frame.

Geographic researchers are on average familiar with multiple reproducible research practices ($\overline{f}=3.26$, Table \ref{tab:overview}), but reported limited use of these practices in their own work ($\overline{p}=1.44$).
More than half of all researchers reported familiarity with with data sharing (86.7\%), open source software (85.3\%), field and lab notebooks (67.0\%), and code sharing (59.2\% ).
However, while familiar with these practice, a far smaller number of researchers reported using these practices regularly in their own work. 
Less than half of the researchers surveyed reported sharing their data (44.5\%), using open source software (38.1\%), or using field or lab notebooks to record their work (40.0\%), or sharing their code (18.8\%) most or all of the time. 
Only a small subset of researchers reported familiarity with the pre-registration of research designs and protocols (27.5\%), or regular use of this practice (2.7\%).

Researcher familiarity and use of reproducible research practices also varied by disciplinary subfield and methodological approach. 
Researchers that identified as physical geographers ($\overline{f_{PH}}=3.71$) or methodologists and GIScientists ($\overline{f_{MT}}=3.97$) reported being familiar with more reproducible researcher practices than human  geographers ($\overline{f_{HU}}=2.35$) and those focused on nature and society ($\overline{f_{NS}}=2.28$).
Researcher practices similarly diverged by subfield, but no subset of researchers reported using on average more than two of these practices regularly in their work.  
When grouped by methodological approach, quantitative and mixed-methods researchers reported substantially greater familiarity ($\overline{f_{QN}}=3.75$, $\overline{f_{MX}}=3.56$) and use ($\overline{p_{QN}}=1.63$, $\overline{p_{MX}}=1.69$) of reproducible researcher practices than qualitative researchers ($\overline{f_{QL}}=1.75$, $\overline{p_{QL}}=0.48$).

Differences in researcher familiarity and use of specific reproducible research practices across subfields and approaches were greatest for practices more typical of quantitative workflows. 
For example, quantitiative and mixed-methods researchers reported familiarity with open source software (QT=96.7\%, MX=89.5\%) and code sharing (QT=81.5\%, MX=55.7\%).
In contrast, a smaller percentage of qualitative researchers reported familiarity with these same practices (OS=52.5\%, Code=12.5\%).
However, even among quantitiative and mixed-methods researchers familiarity with reproducible research practices did not translate into regular use of those practices.
Just over half of quantitative (51.2\%) and mixed-methods (51.8\%) researchers reported sharing their data most of the time.
Less than a third of these same researchers reported sharing their code most of the time. 
These rates were lower for qualitative geographers, of which only a small subset reported using open source software (10.0\%) or field notes (20.0\%), or sharing data (12.5\%) and code (5.0\%) most of the time.

\subsubsection*{Barriers to Reproducible Research}
Aligning with the differences we observed between researcher familiarity with reproducibility enhancing research practices and researcher use of those same practices, respondents to our survey identified numerous factors contributing to a lack or reproducible research across the discipline ($\overline{b}=8.20$). 
Geographic researchers working in different subfields and with different methodological approaches each identified a similar number of barriers to reproducibility (Table \ref{tab:overview}).
Qualitative geographers may be the one group that deviates from this consistent pattern. 
These researchers identified fewer barriers to reproducibility on average, but with a greater variance that left us unable to distinguish this group from any other ($\overline{b_{QL}}=5.97, \sigma_{QL}=3.83$).
To examine differences in researcher beliefs about specific factors potentially hindering reproducibility in the discipline, we divided the 12 factors we examined into three groups --- those related to the research environment, the availability of research artifacts, and study-specific characteristics (Table \ref{tab:barriers}).

\begin{table}[h]
    \centering
    \begin{threeparttable}
    \caption{Barriers to Reproducibility}
    \begin{tabular}{l c c c c c c c c c c}
         \hline
                    & \multicolumn{4}{1}{Subfield}  & & \multicolumn{3}{1}{Approach} & & \\
         Barrier    & HU & PH & NS & MT            & & QN & QL & MX              & & Overall \\
         \hline
         \textit{Research Environment}      & & & & & & & & & & \\
         Pressure to Publish                & 61.2\% & 83.1\% & 54.5\% & 75.0\% & & & & & & 71.5\% \\
         Insufficient oversight             & 56.7\% & 81.6\% & 63.6\% & 82.0\% & & & & & & 71.1\% \\
         Fraud                              & 22.4\% & 32.3\% & 9.1\%  & 39.3\% & & & & & & 28.4\% \\
                                            & & & & & & & & & &\\
         \textit{Artifact Availability}     & & & & & & & & & & \\
         Insufficient metadata              & 62.6\% & 76.9\% & 59.1\% & 78.7\% & & & & & & 80.2\% \\
         Unavailable Data                   & 65.7\% & 89.3\% & 59.1\% & 78.6\% & & & & & & 75.2\% \\
         Unavailable protocol/code          & 59.7\% & 77.0\% & 50.0\% & 86.9\% & & & & & & 71.1\% \\
         Published incomplete results       & 56.7\% & 73.8\% & 63.7\% & 78.7\% & & & & & & 68.4\% \\
         Use restricted data/software       & 44.8\% & 56.9\% & 40.9\% & 82.0\% & & & & & & 57.8\% \\
                                            & & & & & & & & & &\\
         \textit{Study Characteristics}     & & & & & & & & & & \\
         Geographic variation               & 61.2\% & 80.0\% & 81.8\% & 72.2\% & & & & & & 71.5\% \\
         Researcher positionality           & 70.2\% & 50.8\% & 72.8\% & 70.5\% & & & & & & 64.2\% \\
         Chance                             & 55.2\% & 66.2\% & 68.2\% & 65.6\% & & & & & & 62.3\% \\
         Different computation              & 38.8\% & 53.8\% & 31.8\% & 68.9\% & & & & & & 50.9\% \\
                                            & & & & & & & & & &\\
         N                                  & 67 & 65 & 22 & 61 & & 92 & 40 & 85 & & 218 \\
        \hline
    \end{tabular}
    \begin{tablenotes}
        \footnotesize
        \item \textit{HU} Human Geography, \textit{PH} Physical Geography, \textit{NS} Nature and Society, \textit{MT} GIScience and Methods; \textit{QN} Quantitative, \textit{QL} Qualitative, \textit{MX} Mixed Methods. 
        \item Percentage of respondents reporting each factor frequently or occasionally contributed to a lack of reproducibility in geographic research
    \end{tablenotes}
    \label{tab:barriers}
    \end{threeparttable}
\end{table}

Geographic researchers identified the incentive structure of the researcher environment as an important barrier to reproducibility in the discipline.
A majority of geographic researchers identified both the pressure to publish (71.5\%) and insufficient oversight of the research process (71.1\%) as barriers to reproducibility.
A high percentage of physical, methods-focused, and quantitative geographers identified these factors as barriers. 
Qualtiative geographers were least concerned with these factors, but a majority of these researchers still identified both factors as barriers to reproducibility in the field.
Contrary to alarms raised in some of the reproducibility literature, a minority of geographic researchers (28.4\%) believe that the fabrication of data, the manipulation of research results, and similar forms of fraud are a cause of irreproducibility in the discipline.

Researchers identified the unavailability of research artifacts (e.g., data, code) as a barrier to reproducibility, but the importance placed on different artifacts varied by subfield and methodological approach.
A higher percentage physical and methods-focused researchers identified all five of the artifacts we investigated as common barriers to reproducibility than did human and nature-society researchers.
A similar gap existed between qualitative and mixed-methods researchers and qualtiative researchers. 
The largest differences between these groups existed in researchers beliefs about how often the availability of research protocols/code and the use of restricted data or software impacted reproducibility.
More than three-quarters of physical (77.0\%) and methods-focused (86.9\%) geographers identified the availability of protocols/code as contributing to irreproducibility, whereas less than two-thirds of human (59.7\%) and nature-society (50.0\%) identified this same factor. 
Only a large percentage of methods-focused researchers (82.0\%) identified the use of restricted data/software as a common contributor to irreproducibility in the discipline.

A majority of researchers identified geographic variation (71.5\%), researcher positionality (64.2\%), and chance (62.3\%) as study-specific factors limiting the reproducibility of geographic research.
Minor variations in the emphasis placed on these factors exist across subfields and approaches. 
A higher percentage of nature-society (81.8\%) and physical (80.0\%) researchers emphasized the important role the spatial variation and complexity of geographic processes can play when attempting to reproduce geographic research, but this factor was also recognized by researcher across subfields and approaches. 
A smaller percentage of physical geographers placed less emphasis on the impact of researcher positionality may have on reproducibility (50.8\%) when compared to all other subfields.  
Differences between the computational environment used to conduct and original study and a reproduction attempt was generally not seen as a common factor contributing to a lack of reproducibility in the discipline. 
The exception being methods-focused (68.9\%) and quantitative researchers (\%), which aligns with the research practices used in these areas of research.

@@

WHY DID RESEARCHERS ATTEMPT REPRODUCTION - Q10

ACCESS MATERIALS - Q12.
The majority of researchers (82/102) that attempted reproductions were able to access at least some of the data used in the original study, but only (27/102) could access all of the original data.

OUTCOME - Q11 AND OUTCOME BY ACCESS Q12
Geographic researchers report being able to recreate some the results they attempt to reproduce, but few researchers report being able to identically recreate all results (20/102).
